---
title: "Regression Models - "
author: "Jeroen Zonneveld, June 2015"
output: pdf_document
---
##Executive Summary

*Motor Trend*, an automobile trend magazine is interested in exploring the relationship between a set of variables and miles per gallon (MPG) outcome. In this project, we will analyze the `mtcars` dataset from the 1974 Motor Trend US magazine to answer the following questions:

* Is an automatic or manual transmission better for MPG?   
*Manual transmissions achieve a higher value of MPG compared to automatic transmission.*

* Quantify the MPG difference between automatic and manual transmissions.   
*This increase is approximately 1.8 MPG when switching from an automatic transmission to a manual one, with all other variables held constant.*

```{r packages, message = FALSE, warning = FALSE}
library(caret)
```

##Exploratory Data Analysis

The data was extracted from the 1974 Motor Trend US magazine, and comprises 11 aspects of automobile design and performance for 32 automobiles (1973-1974 models). For a description of the data set please folow this [link](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html).

```{r data load}
data(mtcars)
mtcarsRAW   <- mtcars
mtcars$am   <- factor(mtcars$am,labels=c("automatic","manual"))
mtcars$gear <- factor(mtcars$gear)
mtcars$cyl  <- factor(mtcars$cyl)
mtcars$carb <- factor(mtcars$carb)
mtcars$vs   <- factor(mtcars$vs)
```

To explore the relationship of the variable under study, `mpg`, with the other variables in the dataset, we plot `mpg` against each of the other variables (see appendix, figure 1). We try to find the highly correlated variables (>0.75), so we can excluded them from the model.

```{r correlation}
mtcarsCorrelation=cor(mtcarsRAW)
corIndices = findCorrelation(mtcarsCorrelation, cutoff = 0.75); print(corIndices)
```
So the variables `cyl`, `disp`, `mpg` and `am` are highly correlated. As we are interested in the effect of `am` on `mpg` they will be part of the model, we will exclude `cyl` and `disp`. We will use the step function for the selection process.

Figure 2 in th appendix shows a first comparison of the effect of the type of transmission (`am`) on the fuel consumption (`mpg`), 

##Regression Model
First, we build a base regression model that includes all variables as predictors for `mpg`. Next we use the step function to refine our linear model, by selecting the most appropriate variables as predictors, and eliminating those that do not.
```{r models, results='hide'}
baseModel <- lm(mpg ~ ., data = mtcars)
bestModel <- step(baseModel, direction = "both")
```

As we can see in the summary below, the algorithm has selected `cyl`, `hp`, `wt` and `am` as predictors. To check residuals for normality and homoskedasticity, we plot the residuals (see appendix, figure 3). We see that the residuals are normally distributed and homoskedastic.

```{r modelsResult}
summary(bestModel)
```

##Conclusions
bestModel, has successfully explained more than 84% of the variablilty of the data. The model coefficients provide the following insights:

* Miles per gallon increases by a factor of 1.8 (1.8) with manual transmission.
* Miles per gallon decreases by a factor of 0.03 (-0.03) as horspower increases.
* Miles per gallon decreases by a factor of 2.5 (-2.5) for every increase of 1000 lb in  weight.
* Miles per gallon decreases by a factor of 3.03 (-3.03) for 6 cylinders and by a factor of 2.16 (-2.16) for 8 cylinders.
* The intercept is at 33.7 mpg.

The overall p-value is very small (1.506 x 10 ^ -10) which means that there is a very small uncertainty for the sign of the coefficients.


##Appendix

###Figure 1 - Relationships of the different variables

```{r figure1, echo=FALSE}
pairs(mpg ~ ., data = mtcars)
```

###Figure 2 - MPG vs. Transmission Type

```{r figure2, echo=FALSE}
boxplot(mpg ~ am, data = mtcars, col = (c("red","blue")), ylab = "Miles Per Gallon", xlab = "Transmission Type")
```

###Figure 3 - Residuals

```{r figure3, echo=FALSE}
par(mfrow = c(2,2))
plot(bestModel)
```